Nail: A practical interface generator for data formats

### 2014-05-01 paper/osdi14/design.tex
Initial syntax table

\paragraph{Semantic Bijecction.}
Second, in order to generate output, Nail maintaining
a \emph{semantic bijection} between the external format and the internal model. That is, when
generating the corresponding output to a parser input, the two byte streams will have the same
meaning, i.e. be interpreted the same by Nail, but might not be identical.
 A bijection in the traditional sense often does not make sense for data
formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
that tolerates arbitrarily long padding.
%\footnote{Say, the physical layer of most communication
%  protocols is a possibly infinite sequence of symbols that are syntactically nil followed by a
%  pre-determined synchronization sequence and the actual contents of the transmission.}
 Program semantics should be independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded fields \emph{constants}.
Similarly, Nail does not preserve the layout of objects referred to by their offsets. If the grammar
consists of a simple protocol without offset fields, constants and the like, there is a conventional
bijection between  internal models and valid parser inputs.
\paragraph{Hide redundant information.}
Nail's internal model is designed to hide unneeded and redundant information from the application.
Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
and needs to be kept as additional state during parsing. Dependent fields are for example used to represent
lengths, offsets and checksums.

@@

\XXX[In the rest of this section, we first present the design of the Nail language and then show how
Nail's features match these goals.]
\subsection{Basics}
Nail grammars are a set of rules. Each rule assigns a parser, which can be a reference to another
rule, a trivial parser describing an integer of a certain width or a combinator that combines
multiple parsers in some way. We provide the most common combinators familiar from other parser
combinator libraries, such as Parsec\cite{LeijenMeijer:parsec} and Hammer\cite{hammer-parser}.
However, we extend the sequence and choice combinators normally found in parser combinator libraries
with \emph{semantic labels} that name fields in the internal models. The syntax described by a Nail
grammar does not depend on these semantic labels.

\paragraph{Integers and Constraints}
\begin{figure}[tb]
\begin{tabular}{@{}ll@{}}
\toprule
\bf Syntax & \bf Semantics \\
\midrule
\cc{int32} & 32-bit signed integer \\
\multirow{2}{*}{\cc{uint4}}
  & 4-bit unsigned integer, \\
  & returned as an 8-bit value \\
\multirow{2}{*}{\cc{uint8 | 1..3}}
  & 8-bit unsigned integer, \\
  & $1\leq x \leq 3$ \\
\multirow{2}{*}{\cc{uint16 | [1,2]}}
  & unsigned 16-bit integer, \\
  & $x \in \{1,2\}$ \\
\multirow{2}{*}{\cc{int32 | [1,5..255,512]}}
  & signed 32 bit integer, \\
  & $x \in \{ 1, 512 \} \vee 5\leq x\leq 255$ \\
\bottomrule
\end{tabular}
\caption{Example Nail grammars for integer values.}
\label{fig:range}
\end{figure}

Nail's fundamental parsers recognise signed or unsigned with arbitrary lengths up to 64 bits, an
approach popularised by the Hammer framework.\cite{hammer-parser}  
Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
data offset within the TCP header.  In the internal representation, integer values are represented
by the nearest usable integer size, for example a 24-bit unsigned integer will be stored in an
unsigned 32-bit double word.

@@

\paragraph{References.}
To support recursive structures, we include a reference combinator \cc{*}, which is
syntactically equivalent to directly including the parser, but in the internal model adds a level of
indirection to the value. This allows recursive structures to be parsed, for example, balanced
parentheses could be parsed as follows: \cc{paren = < uint8 ="("; optional *paren ;uint8=")">}

@@

If two options the same string could be parser by two options, generated output for the latter
option is not necessarily understood identically by the parser. However, actual data formats are
usually not ambiguous in this sense.
For example, Figure~\ref{fig:choice} demonstrates a simple choice
combinator.

\begin{figure}[h!]
\smaller[0.5]
\begin{verbatim}
  choose {
    A = uint8 | 1..8
    B = uint16 | ..256
  }
\end{verbatim}
\caption{A simple choice combinator that parses either an 8-bit unsigned
integer with a value between 1 and 8 (option A), or a 16-bit unsigned
integer with a value of at most 256 (option B).}
\label{fig:choice}
\end{figure}

@@

\subsection{Dependent fields}

Data formats often contain values that are determined by other values or the layout of information,
such as checksums, duplicated information  or offset and  length fields.
We represent such as values with \emph{dependent fields} and handle them transparently during
parsing and generation without exposing them to the internal model.

@@

\subsection{Input streams and Transformations.}

In order to handle complicated input functions, Nail allows the programmer to
break out of Nail's generated code in a controlled fashion, by manipulating dependent fields and
streams. 
By default, Nail implements streams with in-memory buffers, but the programmer can override this implementation.
While parsing or generating a Nail grammar, there is always an implicit \emph{current stream} and
referred to as \cc{\$current} in the Nail language.

>>>>

\paragraph{Semantic Bijecction.}
Second, in order to generate output, Nail maintaining
a \emph{semantic bijection} between the external format and the internal model. That is, when
generating the corresponding output to a parser input, the two byte streams will have the same
meaning, i.e. be interpreted equivalently by Nail, but might not be identical.
 A bijection in the traditional sense often does not make sense for data
formats. Consider a grammar for a text language that tolerates white space, or a binary protocol
that tolerates arbitrarily long padding.
%\footnote{Say, the physical layer of most communication
%  protocols is a possibly infinite sequence of symbols that are syntactically nil followed by a
%  pre-determined synchronization sequence and the actual contents of the transmission.}
 Program semantics should be independent of the number of padding elements in the input, and Nail therefore
does not expose that information to the programmer. We call such discarded fields \emph{constants}.
Similarly, Nail does not preserve the layout of objects referred to by their offsets. If the grammar
consists of a simple protocol without offset fields, constants and the like, there is a conventional
bijection between  internal models and valid parser inputs.
\paragraph{Hide redundant information.}
Nail's internal model is designed to hide unneeded and redundant information from the application.
Nail introduces \emph{dependent fields}, which contain data that can be computed during generation
and needs to be kept as additional state during parsing. Dependent fields are for example used to represent
lengths, offsets and checksums.

@@

In the rest of this section, we describe the design on the Nail grammar language.
\input{fig-syntax}
\subsection{Basics}
A Nail parser defines both the structure of some external format and a data type to represent that
format. Parsers are constructed by combinators over simpler parsers, an approach popularised by the Parsec
framework\cite{LeijenMeijer:parsec}. We provide the most common combinators familiar from other parser
 combinator libraries, such as Parsec and Hammer\cite{hammer-parser} and extend them so they also
 describe a data type. 
\paragraph{Integers and Constraints}

Nail's fundamental parsers represent signed or unsigned with arbitrary lengths up to 64 bits, an
approach popularised by the Hammer framework.\cite{hammer-parser}  
Note that is possible to define parsers for sub-byte lengths, e.g. to parse the 4-bit
data offset within the TCP header.  In the internal representation, integer values are represented
by the nearest usable integer size, for example a 24-bit unsigned integer will be stored in an
unsigned 32-bit double word.

@@

@@

If two options the same string could be parser by two options, generated output for the latter
option is not necessarily understood identically by the parser. However, actual data formats are
usually not ambiguous in this sense.
For example, Figure~\ref{fig:syntax} demonstrates a simple choice
combinator.

@@

\paragraph{Rules.}
A Nail grammar file consists of rules that assign a parser to a name. Rules are written as
assignments, such as \texttt{ints = many uint8}, which defines a rule called \texttt{ints} with the
body \texttt{many uint8}.  As we will describe in Sections~\ref{s:dependent}
and~\ref{s:transforms}, rules can optionally consume parameters. 
Rules can be invoked in a Nail grammar anywhere a parser can appear. Rule invocations act as though
the body of the rule had been substituted in the code. If parameters appear, they are passed by reference.

Rules allow for recursive grammars. To support recursive data types, we allow introduce the
reference combinator \cc{*}  that does not change the syntax of the external format described, but
introduces a layer of indirection, such as a reference or pointer, to the model data type.

% To support recursive structures, we include a reference combinator \cc{*}, which
%   is syntactically equivalent to directly including the parser, but in the internal model adds a
%   level of indirection to the value. This allows recursive structures to be parsed, for example,
%   balanced parentheses could be parsed as follows: \cc{paren = < uint8 ="("; optional *paren
%     ;uint8=")">}

\subsection{Dependent fields}
\label{s:dependent}
Data formats often contain values that are determined by other values or the layout of information,
such as checksums, duplicated information  or offset and  length fields.
We represent such as values with \emph{dependent fields} and handle them transparently during
parsing and generation without exposing them to the internal model.

@@

\subsection{Input streams and Transformations.}
\label{s:transforms}
In order to handle complicated input functions, Nail allows the programmer to
break out of Nail's generated code in a controlled fashion, by manipulating dependent fields and
streams. 
By default, Nail implements streams with in-memory buffers, but the programmer can override this implementation.
While parsing or generating a Nail grammar, there is always an implicit \emph{current stream} and
referred to as \cc{\$current} in the Nail language.

### 2014-05-01 paper/osdi14/eval.tex
Initial syntax table

\subsection{Data formats}
\label{s:eval-formats}
To answer the first question, we implemented grammars for four protocols with particularly
challenging features and show that Nail is a good fit for these data formats and some aspects of
these formats would be very hard to parse with other parser generators. The following table lists
our example grammars, along with any features that make those protocols hard or impossible to handle
with existing grammar languages and the size of the grammar and its associated transform functions
measured in lines of code.

\begin{tabular}{l|l|l}
\textbf{Protocol} & \textbf{LOC} & \textbf{Problematic features} \\ \hline
DNS messages &  48+64 & Label compression\\
& & Count fields \\
ZIP archives & 92 + 78 & Checksums, \\ 
& & compression, offsets, \\
& &  variable length trailer\\ 
Ethernet frames & 16 & \\
ARP &  10  & \\
IP &  25 & Total Length field, options \\
UDP & 7 & Checksum, Length field \\
ICMP & 5 & Checksum \\

\end{tabular}

\subsubsection{DNS.}

@@

\subsubsection{Ethernet}
\XXX[TODO]
\subsubsection{ZIP Files.}
\input{fig-zip}
\label{s:eval-format-zip}
Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

@@

\noindent\begin{tabular}{l | l | l }
\textbf{System} & \textbf{Nail LoC}  & \textbf{LoC w/o Nail} \\
DNS server & 183 + 48 + 64 & 683 (Hammer parser)\\
&&  ~10.000 (DJBDNS)\\ 
DNS \texttt{host} tool & 98 + 48 + 64  & \XXX[TODO]\\
\texttt{unzip} & 50 + 92 + 78 & 1600 (Info-Zip) \\

\end{tabular}
%%DNS: 
\paragraph{DNS.}

@@

\paragraph{Network stack.}
\XXX[Write this up]
\subsection{Security.} 
We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
analyse a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
immune against those vulnerability classes. Second, we present the results of fuzz-testing our DNS
server.

@@

\XXX[Add fuzzing results - no crashes]
\subsection{Performance.}
\input{fig-perf-dns}
To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
benchmarked our DNS server against BIND.

>>>>

\subsection{Data formats}
\label{s:eval-formats}
To answer the first question, we implemented grammars for four protocols with particularly
challenging features and show that Nail is a good fit for these data formats and some aspects of
these formats would be very hard to parse with other parser generators.
Figure~\ref{fig:eval-protocols} lists various example protocols we have implemented.

\begin{figure}[tb]
\label{fig:eval-protocols}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Protocol} & \textbf{LoC} & \textbf{Problematic features} \\ 
\midrule
DNS messages &  48+64 & Label compression\\
& & Count fields \\
ZIP archives & 92 + 78 & Checksums, \\ 
& & compression, offsets, \\
& &  variable length trailer\\ 
Ethernet frames & 16 & \\
ARP &  10  & \\
IP &  25 & Total Length field, options \\
UDP & 7 & Checksum, Length field \\
ICMP & 5 & Checksum \\
\bottomrule

\end{tabular}
\caption{Example Nail grammars, their code size of the grammar plus  the transformations they use
  and protocol features that cannot be expressed in existing grammar languages. }
\end{figure}
\subsubsection{DNS.}

@@

\subsubsection{ZIP Files.}
\input{fig-zip}
\label{s:eval-format-zip}
Another particularly tricky file format is the ZIP compressed archive format\cite{pkzip}.
ZIP files are normally parsed end-to-beginning. At the end of each ZIP file is an \emph{end-of-directory
header}. This header contains a variable length comment, so it has to be located by scanning
backwards from the end of the file until a magic number and a valid length field is found. Many ZIP
implementations disagree on how to find this header in confusing situations, such as when the
comment contains the valid magic bytes\cite{wolf-berlinsides-zip}.
This end-of-directory header contains the offset and size of the \emph{ZIP directory}, which is an
array of one \emph{directory entry header} per file in the archive.
Each entry stores file meta data, such as file name, compressed and uncompressed size and a checksum,
in addition to the offset of a \emph{local file header}. The local file header duplicates most
information from the directory entry header and the compressed file contents follow it immediately.

@@

\noindent\begin{tabular}{@{}lll@{}} 
\toprule
\textbf{System} & \textbf{Nail LoC}  & \textbf{LoC w/o Nail} \\
\midrule
DNS server & 295 & 683 (Hammer parser)\\
&&  ~10.000 (DJBDNS)\\ 
DNS \texttt{host} tool & 210  & \XXX[TODO]\\
\texttt{unzip} & 220 & 1600 (Info-Zip) \\
\bottomrule
\end{tabular}
%%DNS: 
\paragraph{DNS.}

@@

\subsection{Security.} 
We use a twofold approach to evaluate the security of applications implemented with Nail. First, we
analyse a list of CVE's related to the ZIP file format and argue how our ZIP tools based on Nail are
immune against those vulnerability classes. Second, we present the results of fuzz-testing our DNS
server.

@@

To provide additional assurance that Nail parsers are free of memory corruption attacks, we ran the
DNS fuzzer provided with the Metasploit\cite{mspframework} framework on our server, which sent
randomly corrupted DNS queries to our server for 4 hours, during which it did not crash or trigger
the stack or heap corruption detector.
\subsection{Performance.}
\input{fig-perf-dns}
To estimate the performance impact of using Nail grammars as opposed to hand-written grammars, we
benchmarked our DNS server against BIND.

### 2014-05-01 paper/osdi14/intro.tex
Initial syntax table

Third, many data formats contain redundancies, such as repeating information in multiple structures.
Applications usually do not explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft input that causes inconsistencies.
Furthermore, security vulnerability can occur when an application assumes two repetitions of the
same data to be consistent, such as allocating a buffer based on one size field and copying into
that buffer based on another.

@@

Second, this well-defined internal representation allows Nail to generate the external format from
the internal representation, without requiring the programmer to write additional code.

>>>>

Third, many data formats contain redundancies, such as repeating information in multiple structures.
Applications usually do not explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft input that causes inconsistencies.
Furthermore, security vulnerability can occur when an application assumes two repetitions of the
same data to be consistent, such as allocating a buffer based on the value of one size field and copying into
that buffer based on the value of another.

@@

Second, this well-defined internal representation allows Nail to generate output from
the internal representation, without requiring the programmer to write additional code.

### 2014-05-01 paper/osdi14/relwk.tex
Initial syntax table

Generating parsers and generators from an executable specification
is the core concept of Interface Generators, e.g. in CORBA or more
recently~\cite{varda2008}. However, while interface generators work
very well for existing grammars, they do not allow full control over the
format of the output, so cannot be used to implement legacy protocols.
Very related work has been done at Bell Labs with the PacketTypes
system~\cite{mccann2000packet}, however PacketTypes works only as a
parser, not as an output generator and does not support the expressive
power of parsing expression grammars (PEGs), but rather implements a
C-like structure model enhanced with data-dependent length fields and
constraints.

@@

A notable exception is the Mongrel
web server\footnote{\url{http://mongrel2.org/}} which uses a grammar
for HTTP written in the Ragel~\cite{ragel-paper} regular expression
language. Mongrel was re-written from scratch multiple times to achieve
better scalability and design, yet the grammar could be  re-used across
all iterations~\cite{patterson-citation}.

@@

Input processing vulnerabilities fall into two broad classes.  The first
class is memory safety bugs, such as buffer overflows, which allow an
adversary to corrupt the application's memory using specially crafted
inputs.  These mistakes arise in lower-level languages that do not
provide memory safety guarantees (such as C), and can be partially
mitigated by a wide range of techniques, such as static analysis,
dynamic instrumentation, and address space layout randomization, that
make it more difficult for an adversary to exploit these bugs.
Nail helps developers of lower-level languages avoid these bugs
in the first place.

>>>>

Generating parsers and generators from an executable specification
is the core concept of Interface Generators, e.g. in CORBA or more
recently~\cite{varda2008}. However, while interface generators work
very well for existing grammars, they do not allow full control over the
format of the output, so cannot be used to implement legacy protocols.
Very related work has been done at Bell Labs with the PacketTypes
system~\cite{mccann2000packet}, however PacketTypes works only as a
parser, not as an output generator and does not support the expressive
power of parsing expression grammars (PEGs), but rather implements a
C-like structure model enhanced with  length fields and
constraints. PacketTypes cannot handle complicated encodings such as  compressed data.

@@

A notable exception is the Mongrel
web server~\cite{mongrel} which uses a grammar
for HTTP written in the Ragel~\cite{ragel-paper} regular expression
language. Mongrel was re-written from scratch multiple times to achieve
better scalability and design, yet the grammar could be  re-used across
all iterations~\cite{patterson-citation}.

@@

Input processing vulnerabilities fall into two broad classes.  The first
class is memory safety bugs, such as buffer overflows, which allow an
adversary to corrupt the application's memory using specially crafted
inputs.  These mistakes arise in lower-level languages that do not
provide memory safety guarantees such as C, and can be partially
mitigated by a wide range of techniques, for example static analysis,
dynamic instrumentation, and address space layout randomization, that
make it more difficult for an adversary to exploit these bugs.
Nail helps developers of lower-level languages avoid these bugs
in the first place.
